---
title: "p8105_hw1_jl5297"
author: "Jun Lu"
date: "9/13/2018"
output: github_document
---
## Executive Summary 
This my first-week homework for Data Science, including my solutions to Problems 1 and 2. And I'm praciticing using R Markdown to write reproducible reports, GitHub for version control, and R Projects to organize your work through this homework.

## Load Packages
```{r, message = FALSE, warning = FALSE}
library(tidyverse)
```



## Problem 1
### 1. Create a dataframe
Create a data frame comprised of:  

*   A random sample of size 10 from a uniform[0, 5] distribution
*   A logical vector indicating whether elements of the sample are greater than 2
*   A (length-10) character vector
*   A (length-10) factor vector

```{r p1_dataset}
set.seed(1)
problem1_df = tibble(
    vector_numeric = runif(10, 0, 5),
    vector_logical = vector_numeric > 2,
    vector_character = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    vector_factor = factor(c("male", "female", "male", "male", "male", "male", "male",
                             "male", "male", "female"))
)
```
      
          
### 2. Take the mean
Take the mean of each variable in the dataframe.
```{r take_mean}
mean(problem1_df$vector_numeric)
mean(problem1_df$vector_logical)
mean(problem1_df$vector_character)
mean(problem1_df$vector_factor)
```
It shows that only numeric and logical variables work. In factor and character, NA is returned. Because taking the mean only can apply to numbers. And a logical variable can be interptred into number, false as 0 and true as 1.
     
### 3. Convert variables 
Convert other variables into numeric.
```{r convert1, results = 'hide', warning = FALSE}
as.numeric(problem1_df$vector_logical)
as.numeric(problem1_df$vector_character)
as.numeric(problem1_df$vector_factor)
```
Logical and factor variables can easily convert into numeric.    
Logical: False become 0 and true become 1.  
Factor: Factors convert into integers according to their levels.    
As for character, character variables containing only a number can be converted into numeric.  
     
         
Convert character variable from character to factor to numeric. Convert factor variable from factor to character to numeric.
```{r convert2}
as.numeric(as.factor(problem1_df$vector_character))
as.numeric(as.character(problem1_df$vector_factor))
```
When converting character variable from character to factor to numeric, it can be turned into a integer. When converting factor variable from factor to character to numeric, NA introduced by coercion.
        
## Problem 2
### 1. Create a dataframe
Create a data frame comprised of:
    
* x: a random sample of size 1000 from a standard Normal distribution
* y: a random sample of size 1000 from a standard Normal distribution
* A logical vector indicating whether the x + y > 0
* A numeric vector created by coercing the above logical vector
* A factor vector created by coercing the above logical vector  

```{r p2_dataset}
# I have set the seed in chunk p1_dataset
problem2_df = tibble(
    x = rnorm(1000),
    y = rnorm(1000),
    xy_logical = x + y > 0,
    xy_numeric = as.numeric(xy_logical),
    xy_factor = as.factor(xy_logical)
)
```

* The dataset has  `r ncol(problem2_df)` columns and `r nrow(problem2_df)` rows.
* The mean of x is `r mean(problem2_df$x)` and median of x is `r median(problem2_df$x)`.
* The proportion of cases for which the logical vector is TRUE is `r mean(problem2_df$xy_logical)`.

### 2. Make scatterplots
Make a scatterplot y vs x; color points using the logical variable.
```{r plot1, message = FALSE}
ggplot(problem2_df, aes(x = x, y = y, color = xy_logical)) + geom_point() +
    scale_color_discrete(name = "x+y>0 Logical")
ggsave("plot1.png")
```
There are only two colors, red and blue, corresponding to two logical variables. And red points and blue points are separated by the line of y = -x.

Make a scatterplot y vs x; color points using the numeric variable.
```{r plot2, message = FALSE}
ggplot(problem2_df, aes(x = x, y = y, color = xy_numeric)) + geom_point() + 
    scale_color_continuous(name = "x+y>0 Numeric")
ggsave("plot2.png")
```
The color scale varies between light blue to dark blue. And dark blue points and light blue points are separated by the line of y = -x.

Make a scatterplot y vs x; color points using the factor variable.
```{r plot3, message = FALSE}
ggplot(problem2_df, aes(x = x, y = y, color = xy_factor)) + geom_point() + 
    scale_color_discrete(name = "x+y>0 Factor")
ggsave("plot3.png")
```
There are only two colors, red and blue, corresponding to two factor variables. And red points and blue points are separated by the line of y = -x.



